---
title: Query function usage by package dependencies
output: github_document
params:
  package:
    label: "Package:"
    value: glue
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

```{r, echo = FALSE}
pkg <- params$package

Sys.setenv(PKG = params$package)
write_lines(str_glue("{devtools::revdep(pkg)}_"), str_glue("{pkg}_revdeps"))
```

```{bash, echo = FALSE}
# update CRAN mirror
rsync -q -t --delete -d --include='*tar.gz' cran.r-project.org::CRAN/src/contrib/ cran

# Delete the broken symlinks
find cran -type l -exec test ! -e {} \; -delete

# Construct search patterns

echo "$PKG::[[:alnum:]_.]+
importFrom\(${PKG}[, ]+[[:alnum:]_.]+\)
import\($PKG\)" > search

# Search for function usage in all reverse dependencies
find -L cran -name '*tar.gz' |
  grep -Ff ${PKG}_revdeps |
  xargs -P 8 -n 1 -J % \
    gtar xf % --wildcards '*NAMESPACE' '*/R/*' \
      --to-command 'egrep -o -H --label="$TAR_FILENAME" -f search; true' \
  > ${PKG}_usage ; true
```

```{r, echo = FALSE}
usage <- read_lines(str_glue("{pkg}_usage"))

imports <- grep("NAMESPACE:import", usage, value = TRUE)

raw_calls <- grep("NAMESPACE:import", usage, value = TRUE, invert = TRUE)

calls <-
  as_tibble(raw_calls) %>%
    extract(value, c("file", "fun"), regex = "(.*R)?:?\\w+::(.*)") %>%
    fill(file) %>%
    extract(file, c("pkg", "file"), regex = "(\\w+)/(.*)")
```

# Call analysis

The following counts are all from explicit calls to your package, e.g. `pkg::foo()`.

## Packages with most calls
These are generally the packages which are the heaviest users of your package.

```{r}
calls %>% count(pkg) %>% arrange(desc(n))
```

## Functions most called
There are the functions from your package dependencies are using most frequently.

```{r}
calls %>% 
  count(fun) %>% 
  mutate(percent = scales::percent(n / sum(n))) %>%
  arrange(desc(n)) %>% 
  head(20)
```

## How many packages use each function
This helps determine how broad function usage is across packages.
```{r}
calls %>%
  select(pkg, fun) %>%
  unique() %>%
  count(fun) %>%
  mutate(percent = scales::percent(n / sum(n))) %>%
  arrange(desc(n)) %>%
  head(20)
```

# Imports

The following counts come from dependencies which explicitly import functions from your package with `importFrom()` or `import()`.  While we don't see how often they are using each function in this case, we can see which functions are being imported.

```{r, echo = FALSE}
raw_imports <- grep("NAMESPACE:import", usage, value = TRUE)

full_imports <- grep("NAMESPACE:import\\(", raw_imports, value = TRUE) %>%
  sub("/NAMESPACE.*", "", .)
```

## Packages with full imports

`r length(full_imports)` have 'full' imports, with `import(pkg)` in their NAMESPACE. It is difficult to determine function usage of these packages.

```{r}
full_imports
```

```{r, echo = FALSE}
selective_imports <-
  grep("importFrom", imports, value = TRUE) %>%
    as_tibble() %>%
    extract(value, c("pkg", "fun"), "^(\\w+).*,([[:alpha:]._]+)\\)$")
```

## Pkgs with most functions imported
```{r}
selective_imports %>% count(pkg) %>% arrange(desc(n))
```

## Which functions are most often imported?

These are the functions which your dependencies find most useful.

```{r}
selective_imports %>%
  count(fun) %>%
  mutate(percent = scales::percent(n / sum(n))) %>%
  arrange(desc(n)) %>%
  head(20)
```

## Which functions are never used by dependencies?

These are the functions no dependency is using either by calls or importing. These functions either need better documentation / publicity, are meant for interactive use rather than in packages, or do not provide a useful function and should be considered for removal.

```{r}
exports <- getNamespaceExports(pkg)

exports[
  !exports %in% c(calls$fun, selective_imports$fun)
]
```
